---
title: '[WWS 586A]: Problem Set 2'
author: "Jason Anastasopoulos"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

### [WWS 586a] Problem Set 2

For this exercise, you will use the R package *twitteR* and *jsonlite* to collect data from Twitter and the GovTrack API and clean the text data.

### Guidelines

All problem sets must be submitted as two files:

1. A *R Markdown* file with the following format: "lastname_firstname_ps2.Rmd". Ie. for me this would be: "anastasopoulos_jason_ps2.Rmd"

2. A compiled *R Markdown* file in HTML with the following format: "lastname_firstname_ps2.html" Ie. for me this would be: "anastasopoulos_jason_ps2.html"

Please only fill in the sections labelled "YOUR CODE HERE"
  
### Rep. Nancy Pelosi's Tweets

1. Set up OAuth credentials which you can use to access the Twitter database by establishing an account on the [Twitter Developer Page](https://apps.twitter.com)
	
2. Use the *twitteR* package to collect the most recent 100 tweets from *@NancyPelosi* along with: (1) how many times the Tweets were retweeted; (2) the number of likes that each Tweet received; (3) the date and time of the Tweet and; (4) the city of the Tweet (if available).


```{r}
# Let's first load the R packages and the data
library(pacman)

# This loads and installs the packages you need at once
pacman::p_load(foreign,twitteR,jsonlite)


##### YOUR CODE HERE ###################################







##### YOUR CODE HERE ###################################
```


3. Build a data frame  with this information and write the data frame to a .csv file labeled ``pelosi-YourLastName.csv.''


```{r}
##### YOUR CODE HERE ###################################




##### YOUR CODE HERE ###################################
```

### Tapping Into APIs Using *jsonlite*

Using the GovTrack API, find bills with the keyword ``cybersecurity'' in them.

1. Build a data frame with the first 20 bills which contains (1) the bill titles; (2) the bill id; (3) the last name of the bill sponsor; (4) the bill sponsor's party.

```{r}
##### YOUR CODE HERE ###################################




##### YOUR CODE HERE ###################################
```

2. Build a data frame with this information and write the data frame to a .csv file labeled ``cybersecurity-YourLastName.csv.''


```{r}
##### YOUR CODE HERE ###################################




##### YOUR CODE HERE ###################################
```

